---
title: "COVID-19 Impact on Helsinki Shared Bikes"
runtime: shiny
output: html_document
author: Muhammad Sarmad Kamal Chaudhry and Yubin Fu
---

```{r setup, include=FALSE, message=FALSE}
library(tidyverse)
library(lubridate)
library(stringr)
library(ggmap)
library(shiny)
library(base)
library(gridExtra)
library(forcats)
library(dplyr)
library(rsconnect)
library(plotly)
```



```{r, include=FALSE}
finland_covid <- read.csv("~/Mscs 264b S21/Submit/Finland Covid .csv", nrows=40)

helsinki_bikes <- read.csv("~/Mscs 264b S21/Submit/Helsinki Bikes.csv", nrows=331556)

```

```{r, include=FALSE}

helsinki_bikes1 <- helsinki_bikes %>%
  mutate(duration..sec.. = duration..sec.. /60) %>%
  separate(col = departure, into  = c('departure_date', 'departure_time'), 
           sep = ' ') %>%
  separate(col = return, into  = c('return_date', 'return_time'),
           sep = ' ') 

```

```{r,  include=FALSE}

helsinki_bikes2 <- helsinki_bikes1%>%
  mutate(departure_date = format(as.Date(departure_date),'%m/%d/%Y'),
         return_date = format(as.Date(return_date),'%m/%d/%Y')) %>%
  rename(air_temp_cels = Air.temperature..degC., 
         avg_speed_km_h = avg_speed..km.h.,
         distance_meter = distance..m.,
         trip_duration_mins = duration..sec..) %>%
  mutate(departure_date = str_replace(departure_date, "/2020", "")) %>%
  select(departure_date, return_date, departure_name, return_name, 
         distance_meter, trip_duration_mins, air_temp_cels, departure_latitude, departure_longitude, return_latitude, return_longitude)
```


```{r,  include=FALSE}
finland_covid1 <- finland_covid%>%
  mutate(date = as.Date(date, "%m/%d/%Y")) %>%
  mutate(date = str_replace_all(date, "-", "/")) %>%
  mutate(date = str_replace(date, "20/", "")) %>%
  rename(departure_date = date) %>%
  select(departure_date, total_cases, positive_rate,
         new_cases, total_deaths, total_deaths, total_tests)


```

```{r, include=FALSE}

final <- merge(finland_covid1, helsinki_bikes2, by = "departure_date")

```

**COVID-19 Positivity Rate, Deaths, and Bike Sharing:**

```{r, include = FALSE}
final1 <- final %>%
  group_by(departure_date)%>%
  summarise(positivity = mean(positive_rate),
            deaths = mean(total_deaths),
            bike_rides = n())
```

```{r, echo=FALSE}
inputPanel(
  selectInput("y1", label = "Y-axis variable:",
              choices = c("COVID-19 Deaths" = "deaths",
                          "Bike Rides" = "bike_rides",
                          "Positivity Rate" = "positivity")),
  sliderInput("size1", label = "Point size",
              min = 1, max = 3, value = 1, step = 0.1),
  sliderInput("alpha1", label = "Point transparency",
              min = 0.2, max = 1, value = 0.25, step = 0.1)
  )
renderPlot({
  ggplot(final1, aes_string(x = "departure_date", y = input$y1))+
    geom_point(size=input$size1, alpha=input$alpha1) +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
})

```
**Bike Rides Analysis by Top 5 Neighborhoods:**

```{r, include = FALSE}
helsinki_bikes3<- helsinki_bikes2 %>%
  group_by(departure_date, departure_name)%>%
  summarise(
            temperature = mean(air_temp_cels),
            distance_average = mean(distance_meter),
            bike_share = n()) %>%
  filter(departure_name == "Arabian kauppakeskus" |
           departure_name == "Itämerentori" |
           departure_name == "Kalasatama (M)" |
           departure_name == "Ympyrätalo" |
           departure_name == "Pasilan asema")
```

```{r, echo=FALSE}

inputPanel(
  selectInput("y2", label = "Options:",
              choices = c("Temperature" = "temperature",
                          "Bike Rides" = "bike_share",
                          "Average Distance (m)" = "distance_average")),
    sliderInput("size1", label = "Point size",
              min = 1, max = 3, value = 1, step = 0.1),
  sliderInput("alpha1", label = "Point transparency",
              min = 0.2, max = 1, value = 0.25, step = 0.1)
  )

renderPlot({
  ggplot(helsinki_bikes3, aes_string(x = "departure_date", y = input$y2, color = "departure_name"))+
    geom_point(size=input$size1, alpha=input$alpha1) +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
})

```


**Maps of Helsinki Bike Sharing Departure and Return:**

These maps are important for our analysis since it shows that the bikes were returned to the exact place they were loaned from. This tells us that COVID-19 was probably less likely to spread in this case as opposed to if the bikes were loaned from one place and left unsanitized in another part of Helsinki. 

```{r, echo=FALSE, message=FALSE}
departure <- qmplot(departure_longitude, departure_latitude, data = helsinki_bikes2,maptype = "toner-lite", color = I("red"))

departure

return <- qmplot(return_longitude, return_latitude, data = helsinki_bikes2,maptype = "toner-lite", color = I("blue"))

return
```

